IML Term project

### 2022-12-07 main_markdown.rmd
NB, LDA, QDA testing.

### PCA
#### PCA for all three versions of data 
```{r PCA}
pcaOrig <- prcomp(npf[,1:100], center = T, scale.=T)
cumsum(pcaOrig$sdev^2)/sum(pcaOrig$sdev^2)
x1 <- as.matrix(npf[,1:100])%*%pcaOrig$rotation[,1:12]
pca.168 <- prcomp(npf.168[,1:36], center = T, scale. = T)
cumsum(pca.168$sdev^2)/sum(pca.168$sdev^2)
x2 <- as.matrix(npf.168[,1:36])%*%pca.168$rotation[,1:12]
pca.sp <- prcomp(npfsp[,1:18], center = T, scale. = T)
cumsum(pca.sp$sdev^2)/sum(pca.sp$sdev^2)
x3 <- as.matrix(npfsp[,1:18])%*%pca.sp$rotation[,1:11]
```

@@

## Testing classifiers as they are

@@

```{r naive bayes}
library(e1071)
nb.fit <- naiveBayes(class2 ~ . , data = npf_train, laplace = 1)
nb.pred.cv <- cv.prob(class2 ~ ., npf, naiveBayes) # Hidas prosessi, tehdään siksi vaan kerran

nb_acc <- data.frame(train = accuracy(predict(nb.fit, newdata = npf_train, type = "raw")[,2], npf_train$class2),
                     test = accuracy(predict(nb.fit, newdata = npf_test, type = "raw")[,2], npf_test$class2),
                     CV = accuracy(nb.pred.cv, npf$class2))
# Kaikkiin tulee Inf
nb_perp <- data.frame(train = perplexity(predict(nb.fit, newdata = npf_train, type = "raw")[,2], npf_train$class2),
                      test = perplexity(predict(nb.fit, newdata = npf_test, type = "raw")[,2], npf_test$class2),
                      CV = perplexity(nb.pred.cv, npf$class2))
kable(cbind(nb_acc,
            nb_perp)) %>%
  add_header_above(c("Accuracy" = 3, "Perplexity" = 3)) %>%
  kable_styling(latex_options = "HOLD_position")
```

### LDA
```{r lda}
library(MASS)
lda.fit <- lda(class2 ~ . , data = npf_train)
# Cross-validation parametrit
n = nrow(npf) # number of rows in the data matrix
k = min(n, 10) # number of cross-validation folds
split = kpart(n, k) # the split of n data items into k folds
lda.pred.cv <- rep(0,n)

# cv
for (i in 1:k) {
  ## go through all folds, train on other folds, and make a prediction
  mod <- lda(class2 ~ ., npf[split != i, ])
  lda.pred.cv[split == i] <- predict(mod, npf[split == i, ])$posterior[,2]
}
lda_acc <- data.frame(train = accuracy(predict(lda.fit, newdata = npf_train)$posterior[,2], npf_train$class2),
                      test = accuracy(predict(lda.fit, newdata = npf_train)$posterior[,2], npf_test$class2),
                      CV = accuracy(lda.pred.cv, npf$class2))
lda_perp <- data.frame(train = perplexity(predict(lda.fit, newdata = npf_train)$posterior[,2], npf_train$class2),
                      test = perplexity(predict(lda.fit, newdata = npf_train)$posterior[,2], npf_test$class2),
                      CV = perplexity(lda.pred.cv, npf$class2))
kable(cbind(lda_acc,
            lda_perp)) %>%
  add_header_above(c("Accuracy" = 3, "Perplexity" = 3))%>%
  kable_styling(latex_options = "HOLD_position")
```

### QDA
```{r qda}
library(MASS)
qda.fit <- qda(class2 ~ . , data = npf_train)
# Cross-validation parametrit
n = nrow(npf) # number of rows in the data matrix
k = min(n, 10) # number of cross-validation folds
split = kpart(n, k) # the split of n data items into k folds
qda.pred.cv <- rep(0,n)

# cv
for (i in 1:k) {
  ## go through all folds, train on other folds, and make a prediction
  mod <- qda(class2 ~ ., npf[split != i, ])
  qda.pred.cv[split == i] <- predict(mod, npf[split == i, ])$posterior[,2]
}
qda_acc <- data.frame(train = accuracy(predict(qda.fit, newdata = npf_train)$posterior[,2], npf_train$class2),
                      test = accuracy(predict(qda.fit, newdata = npf_train)$posterior[,2], npf_test$class2),
                      CV = accuracy(qda.pred.cv, npf$class2))
# Taas Inf :(
qda_perp <- data.frame(train = perplexity(predict(qda.fit, newdata = npf_train)$posterior[,2], npf_train$class2),
                      test = perplexity(predict(qda.fit, newdata = npf_train)$posterior[,2], npf_test$class2),
                      CV = perplexity(qda.pred.cv, npf$class2))
kable(cbind(qda_acc,
            qda_perp)) %>%
  add_header_above(c("Accuracy" = 3, "Perplexity" = 3)) %>%
  kable_styling(latex_options = "HOLD_position")
```

>>>>

### PCA
#### PCA for all three versions of data 
```{r PCA}
pcaOrig <- prcomp(npf[,1:100], center = T, scale.=T)
cumsum(pcaOrig$sdev^2)/sum(pcaOrig$sdev^2)
x1 <- as.matrix(npf[,1:100])%*%pcaOrig$rotation[,1:12]
pca.168 <- prcomp(npf.168[,1:36], center = T, scale. = T)
cumsum(pca.168$sdev^2)/sum(pca.168$sdev^2)
x2 <- as.matrix(npf.168[,1:36])%*%pca.168$rotation[,1:12]
pca.sp <- prcomp(npfsp[,1:18], center = T, scale. = T)
cumsum(pca.sp$sdev^2)/sum(pca.sp$sdev^2)
x3 <- as.matrix(npfsp[,1:18])%*%pca.sp$rotation[,1:11]

npf.pca <- as.data.frame(cbind(x1, npf$class2))
colnames(npf.pca) <- c(colnames(x1), "class2")
npf.168.pca <- as.data.frame(cbind(x2, npf.168$class2))
colnames(npf.168.pca) <- c(colnames(x2), "class2")
```

@@

## Naive Bayes, LDA and QDA on the original data, 168-data and the PCA'd versions of these data sets

@@

```{r naive bayes}
library(e1071)

# Alkuperäinen data
nb.fit <- naiveBayes(class2 ~ . , data = npf_train)
nb.pred.cv <- cv.prob(class2 ~ ., npf, naiveBayes) # Hidas prosessi, tehdään siksi vaan kerran
nb.pred.loocv <- loocv.prob(class2 ~ ., npf, naiveBayes) # Sama

nb_acc <- data.frame(train = accuracy(predict(nb.fit, newdata = npf_train, type = "raw")[,2], npf_train$class2),
                     test = accuracy(predict(nb.fit, newdata = npf_test, type = "raw")[,2], npf_test$class2),
                     CV = accuracy(nb.pred.cv, npf$class2),
                     LOOCV = accuracy(nb.pred.loocv, npf$class2))

nb_perp <- data.frame(train = perplexity(predict(nb.fit, newdata = npf_train, type = "raw")[,2], npf_train$class2),
                      test = perplexity(predict(nb.fit, newdata = npf_test, type = "raw")[,2], npf_test$class2),
                      CV = perplexity(nb.pred.cv, npf$class2),
                      LOOCV = perplexity(nb.pred.loocv, npf$class2))
kable(cbind(nb_acc,
            nb_perp)) %>%
  add_header_above(c("Accuracy" = 4, "Perplexity" = 4)) %>%
  kable_styling(latex_options = "HOLD_position")

# 168-data
nb.pred.cv.168 <- cv.prob(class2 ~ ., npf.168, naiveBayes) # Hidas prosessi, tehdään siksi vaan kerran
nb.pred.loocv.168 <- loocv.prob(class2 ~ ., npf.168, naiveBayes) # Sama

nb_acc.168 <- data.frame(CV = accuracy(nb.pred.cv.168, npf.168$class2),
                     LOOCV = accuracy(nb.pred.loocv.168, npf.168$class2))

nb_perp.168 <- data.frame(CV = perplexity(nb.pred.cv.168, npf.168$class2),
                      LOOCV = perplexity(nb.pred.loocv.168, npf.168$class2))
kable(cbind(nb_acc.168,
            nb_perp.168)) %>%
  add_header_above(c("Accuracy" = 2, "Perplexity" = 2)) %>%
  kable_styling(latex_options = "HOLD_position")

Using PCA on the original and the 16.8m-dataset.

```{r naive bayes pca}

# Alkuperäinen data
nb.pred.cv.pca <- cv.prob(class2 ~ ., npf.pca, naiveBayes) # Hidas prosessi, tehdään siksi vaan kerran
nb.pred.loocv.pca <- loocv.prob(class2 ~ ., npf.pca, naiveBayes) # Sama

nb_acc.pca <- data.frame(CV = accuracy(nb.pred.cv.pca, npf.pca$class2),
                         LOOCV = accuracy(nb.pred.loocv.pca, npf.pca$class2))

nb_perp.pca <- data.frame(CV = perplexity(nb.pred.cv.pca, npf.pca$class2),
                          LOOCV = perplexity(nb.pred.loocv.pca, npf.pca$class2))
kable(cbind(nb_acc.pca,
            nb_perp.pca)) %>%
  add_header_above(c("Accuracy" = 2, "Perplexity" = 2)) %>%
  kable_styling(latex_options = "HOLD_position")

# 168-data
nb.pred.cv.168.pca <- cv.prob(class2 ~ ., npf.168.pca, naiveBayes) # Hidas prosessi, tehdään siksi vaan kerran
nb.pred.loocv.168.pca <- loocv.prob(class2 ~ ., npf.168.pca, naiveBayes) # Sama

nb_acc.168.pca <- data.frame(CV = accuracy(nb.pred.cv.168, npf.168.pca$class2),
                     LOOCV = accuracy(nb.pred.loocv.168, npf.168.pca$class2))

nb_perp.168.pca <- data.frame(CV = perplexity(nb.pred.cv.168, npf.168.pca$class2),
                      LOOCV = perplexity(nb.pred.loocv.168, npf.168.pca$class2))
kable(cbind(nb_acc.168.pca,
            nb_perp.168.pca)) %>%
  add_header_above(c("Accuracy" = 2, "Perplexity" = 2)) %>%
  kable_styling(latex_options = "HOLD_position")

### LDA

On the original dataset and the 16.8m-dataset.
```{r lda, warning=FALSE}
library(MASS)

# Cross-validation parametrit
n = nrow(npf) # number of rows in the data matrix
k = min(n, 10) # number of cross-validation folds
split = kpart(n, k) # the split of n data items into k folds

lda.pred.cv <- rep(0,n)
lda.pred.cv.168 <- rep(0,n)

# cv
for (i in 1:k) {
  ## go through all folds, train on other folds, and make a prediction
  mod <- lda(class2 ~ ., npf[split != i, ])
  lda.pred.cv[split == i] <- predict(mod, npf[split == i, ])$posterior[,2]
  ## 168
  mod <- lda(class2 ~ ., npf.168[split != i, ])
  lda.pred.cv.168[split == i] <- predict(mod, npf.168[split == i, ])$posterior[,2]
}

lda.pred.loocv <- rep(0,n)
lda.pred.loocv.168 <- rep(0,n)

# loocv
for(i in 1:dim(npf)[1]){
  mod <- lda(class2 ~ ., npf[split != i, ])
  lda.pred.loocv[split == i] <- predict(mod, npf[split == i, ])$posterior[,2]
  ## 168
  mod <- lda(class2 ~ ., npf.168[split != i, ])
  lda.pred.loocv.168[split == i] <- predict(mod, npf.168[split == i, ])$posterior[,2]
}

# npf
lda_acc <- data.frame(CV = accuracy(lda.pred.cv, npf$class2),
                      LOOCV = accuracy(lda.pred.loocv, npf$class2))
lda_perp <- data.frame(CV = perplexity(lda.pred.cv, npf$class2),
                       LOOCV = perplexity(lda.pred.loocv, npf$class2))
kable(cbind(lda_acc,
            lda_perp)) %>%
  add_header_above(c("Accuracy" = 2, "Perplexity" = 2))%>%
  kable_styling(latex_options = "HOLD_position")

# 168
lda_acc.168 <- data.frame(CV = accuracy(lda.pred.cv.168, npf$class2),
                          LOOCV = accuracy(lda.pred.loocv.168, npf$class2))
lda_perp.168 <- data.frame(CV = perplexity(lda.pred.cv.168, npf$class2),
                           LOOCV = perplexity(lda.pred.loocv.168, npf$class2))
kable(cbind(lda_acc.168,
            lda_perp.168)) %>%
  add_header_above(c("Accuracy" = 2, "Perplexity" = 2))%>%
  kable_styling(latex_options = "HOLD_position")
```

Using PCA on the original and the 16.8m-dataset.

```{r lda pca, warning=FALSE}

# Cross-validation parametrit
n = nrow(npf) # number of rows in the data matrix
k = min(n, 10) # number of cross-validation folds
split = kpart(n, k) # the split of n data items into k folds

lda.pred.cv.pca <- rep(0,n)
lda.pred.cv.168.pca <- rep(0,n)

# cv
for (i in 1:k) {
  ## go through all folds, train on other folds, and make a prediction
  mod <- lda(class2 ~ ., npf.pca[split != i, ])
  lda.pred.cv.pca[split == i] <- predict(mod, npf.pca[split == i, ])$posterior[,2]
  ## 168
  mod <- lda(class2 ~ ., npf.168.pca[split != i, ])
  lda.pred.cv.168.pca[split == i] <- predict(mod, npf.168.pca[split == i, ])$posterior[,2]
}

lda.pred.loocv.pca <- rep(0,n)
lda.pred.loocv.168.pca <- rep(0,n)

# loocv
for(i in 1:dim(npf)[1]){
  mod <- lda(class2 ~ ., npf.pca[split != i, ])
  lda.pred.loocv.pca[split == i] <- predict(mod, npf.pca[split == i, ])$posterior[,2]
  ## 168
  mod <- lda(class2 ~ ., npf.168.pca[split != i, ])
  lda.pred.loocv.168.pca[split == i] <- predict(mod, npf.168.pca[split == i, ])$posterior[,2]
}

# npf
lda_acc.pca <- data.frame(CV = accuracy(lda.pred.cv.pca, npf.pca$class2),
                          LOOCV = accuracy(lda.pred.loocv.pca, npf.pca$class2))
lda_perp.pca <- data.frame(CV = perplexity(lda.pred.cv.pca, npf.pca$class2),
                           LOOCV = perplexity(lda.pred.loocv.pca, npf.pca$class2))
kable(cbind(lda_acc.pca,
            lda_perp.pca)) %>%
  add_header_above(c("Accuracy" = 2, "Perplexity" = 2))%>%
  kable_styling(latex_options = "HOLD_position")

# 168
lda_acc.168.pca <- data.frame(CV = accuracy(lda.pred.cv.168.pca, npf.168.pca$class2),
                              LOOCV = accuracy(lda.pred.loocv.168.pca, npf.168.pca$class2))
lda_perp.168.pca <- data.frame(CV = perplexity(lda.pred.cv.168.pca, npf.168.pca$class2),
                               LOOCV = perplexity(lda.pred.loocv.168.pca, npf.168.pca$class2))
kable(cbind(lda_acc.168.pca,
            lda_perp.168.pca)) %>%
  add_header_above(c("Accuracy" = 2, "Perplexity" = 2))%>%
  kable_styling(latex_options = "HOLD_position")
```

### QDA
```{r qda, warning=FALSE}
library(MASS)

# Cross-validation parametrit
n = nrow(npf) # number of rows in the data matrix
k = min(n, 10) # number of cross-validation folds
split = kpart(n, k) # the split of n data items into k folds

qda.pred.cv <- rep(0,n)
qda.pred.cv.168 <- rep(0,n)

# cv
for (i in 1:k) {
  ## go through all folds, train on other folds, and make a prediction
  mod <- qda(class2 ~ ., npf[split != i, ])
  qda.pred.cv[split == i] <- predict(mod, npf[split == i, ])$posterior[,2]
  ## 168
  mod <- qda(class2 ~ ., npf.168[split != i, ])
  qda.pred.cv.168[split == i] <- predict(mod, npf.168[split == i, ])$posterior[,2]
}

qda.pred.loocv <- rep(0,n)
qda.pred.loocv.168 <- rep(0,n)

# loocv
for(i in 1:dim(npf)[1]){
  mod <- qda(class2 ~ ., npf[split != i, ])
  qda.pred.loocv[split == i] <- predict(mod, npf[split == i, ])$posterior[,2]
  ## 168
  mod <- qda(class2 ~ ., npf.168[split != i, ])
  qda.pred.loocv.168[split == i] <- predict(mod, npf.168[split == i, ])$posterior[,2]
}

# npf
qda_acc <- data.frame(CV = accuracy(qda.pred.cv, npf$class2),
                      LOOCV = accuracy(qda.pred.loocv, npf$class2))
qda_perp <- data.frame(CV = perplexity(qda.pred.cv, npf$class2),
                       LOOCV = perplexity(qda.pred.loocv, npf$class2))
kable(cbind(qda_acc,
            qda_perp)) %>%
  add_header_above(c("Accuracy" = 2, "Perplexity" = 2))%>%
  kable_styling(latex_options = "HOLD_position")

# 168
qda_acc.168 <- data.frame(CV = accuracy(qda.pred.cv.168, npf$class2),
                          LOOCV = accuracy(qda.pred.loocv.168, npf$class2))
qda_perp.168 <- data.frame(CV = perplexity(qda.pred.cv.168, npf$class2),
                           LOOCV = perplexity(qda.pred.loocv.168, npf$class2))
kable(cbind(qda_acc.168,
            qda_perp.168)) %>%
  add_header_above(c("Accuracy" = 2, "Perplexity" = 2))%>%
  kable_styling(latex_options = "HOLD_position")
```

Using PCA on the original and the 16.8m-dataset.

```{r qda pca, warning=FALSE}

# Cross-validation parametrit
n = nrow(npf) # number of rows in the data matrix
k = min(n, 10) # number of cross-validation folds
split = kpart(n, k) # the split of n data items into k folds

qda.pred.cv.pca <- rep(0,n)
qda.pred.cv.168.pca <- rep(0,n)

# cv
for (i in 1:k) {
  ## go through all folds, train on other folds, and make a prediction
  mod <- qda(class2 ~ ., npf.pca[split != i, ])
  qda.pred.cv.pca[split == i] <- predict(mod, npf.pca[split == i, ])$posterior[,2]
  ## 168
  mod <- qda(class2 ~ ., npf.168.pca[split != i, ])
  qda.pred.cv.168.pca[split == i] <- predict(mod, npf.168.pca[split == i, ])$posterior[,2]
}

qda.pred.loocv.pca <- rep(0,n)
qda.pred.loocv.168.pca <- rep(0,n)

# loocv
for(i in 1:dim(npf)[1]){
  mod <- qda(class2 ~ ., npf.pca[split != i, ])
  qda.pred.loocv.pca[split == i] <- predict(mod, npf.pca[split == i, ])$posterior[,2]
  ## 168
  mod <- qda(class2 ~ ., npf.168.pca[split != i, ])
  qda.pred.loocv.168.pca[split == i] <- predict(mod, npf.168.pca[split == i, ])$posterior[,2]
}

# npf
qda_acc.pca <- data.frame(CV = accuracy(qda.pred.cv.pca, npf.pca$class2),
                          LOOCV = accuracy(qda.pred.loocv.pca, npf.pca$class2))
qda_perp.pca <- data.frame(CV = perplexity(qda.pred.cv.pca, npf.pca$class2),
                           LOOCV = perplexity(qda.pred.loocv.pca, npf.pca$class2))
kable(cbind(qda_acc.pca,
            qda_perp.pca)) %>%
  add_header_above(c("Accuracy" = 2, "Perplexity" = 2))%>%
  kable_styling(latex_options = "HOLD_position")

# 168
qda_acc.168.pca <- data.frame(CV = accuracy(qda.pred.cv.168.pca, npf.168.pca$class2),
                              LOOCV = accuracy(qda.pred.loocv.168.pca, npf.168.pca$class2))
qda_perp.168.pca <- data.frame(CV = perplexity(qda.pred.cv.168.pca, npf.168.pca$class2),
                               LOOCV = perplexity(qda.pred.loocv.168.pca, npf.168.pca$class2))
kable(cbind(qda_acc.168.pca,
            qda_perp.168.pca)) %>%
  add_header_above(c("Accuracy" = 2, "Perplexity" = 2))%>%
  kable_styling(latex_options = "HOLD_position")
```
